DraftFunctions

### 2021-05-07 R/DraftFunctions.Rmd
"addtreedim" function

# Fonction de contrôle & formatage de l’INPUT “data” de la fonction pilot (“datacheckformat”):
INPUT: data = Inventaires botaniques Guyafor/ONF (data frame)
OUTPUT: data = Inventaires botaniques type “Guyafor”(data frame)

vérification du type d’inventaire : 
* Test présence des variables.
* Test class des variables.
Si les tests précédents ont tous reçu =TRUE, STOP de la fonction dataformat. 
Sinon mettre en forme “inventaire Guyafor” (nom des var, leurs classes, unités (mettre des limites avec range),...)
re-vérification du type d’inventaire : 
* Test présence des variables
* Test class des variables
Si les tests précédent ont tous reçu =TRUE, STOP de la fonction dataformat.
Sinon Message : “Votre format de données renseignés à l’argument “data”, n’est pas compatible. Voir la vignette du package pour y trouver le format requis.”

```{r}
data("Paracou6_2016") #data in the environment
lapply(Paracou6_2016, class) #give the class of alla the data variables
```

```{r Create a broken dataset}
#in inst/extdata
BrokenParacou6_2016 <- Paracou6_2016 %>% 
  select(-Species) %>% # missing column
  rename(DBH = CircCorr) %>% # other column name
  rename(PLOT = Plot) %>% # variable name in upper case
  mutate(idTree = as.character(as.numeric(idTree))) %>% #an other variable class
  mutate(Lon = round(Lon, digits =0)) %>% # no digits
  mutate(Genus = toupper(Genus)) # values in upper case
```

@@

'pouet' %in% names(Paracou6_2016)
if (!is.null(Paracou6_2016$pouet)) 
    stop("pouet parameter should be an integer (column name), or NULL if no presence column.")

@@

```{r}
# Variables presence check
"Plot" %in% names(Paracou6_2016)
"CensusYear" %in% names(Paracou6_2016)
"idTree" %in% names(Paracou6_2016)
"Family" %in% names(Paracou6_2016)
"Genus" %in% names(Paracou6_2016)
"Species" %in% names(Paracou6_2016)
"CircCorr" %in% names(Paracou6_2016)
"CodeAlive" %in% names(Paracou6_2016)
"CommercialSp" %in% names(Paracou6_2016)
"UTMZone" %in% names(Paracou6_2016)
'Lat' %in% names(Paracou6_2016)
'Lon' %in% names(Paracou6_2016)
'VernName' %in% names(Paracou6_2016)
'Xfield' %in% names(Paracou6_2016)
'Yfield' %in% names(Paracou6_2016)
'Xutm' %in% names(Paracou6_2016)
'Yutm' %in% names(Paracou6_2016)

@@

```{r}
# Variables class check
 if (!inherits(Paracou6_2016$idTree, "integer") && !is.null(idTree)) #majuscule ?
    stop("idTree parameter should be an integer (column name), or NULL if no presence column.")

if (!inherits(Paracou6_2016$Plot, "character") && !is.null(Plot))
    stop("Plot parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Xfield, "numeric") && !is.null(Xfield))
    stop("Xfield parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Yfield, "numeric") && !is.null(Yfield))
    stop("Yfield parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Xutm, "numeric") && !is.null(Xutm))
    stop("Xutm parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Yutm, "numeric") && !is.null(Yutm))
    stop("Yutm parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$UTMZone, "integer") && !is.null(UTMZone))
    stop("UTMZone parameter should be an integer (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Lat, "numeric") && !is.null(Lat))
    stop("Lat parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Lon, "numeric") && !is.null(Lon))
    stop("Lon parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Family, "character") && !is.null(Family))
    stop("Family parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Genus, "character") && !is.null(Genus))
    stop("Genus parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Species, "character") && !is.null(Species))
    stop("Species parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$VernName, "character") && !is.null(VernName))
    stop("VernName parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$CommercialSp, "logical") && !is.null(CommercialSp))
    stop("CommercialSp parameter should be logical (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$CensusYear, "integer") && !is.null(CensusYear))
    stop("CensusYear parameter should be an integer (column name), or NULL if no presence column.")

  if (!inherits(Paracou6_2016$CodeAlive, "logical") && !is.null(CodeAlive))
    stop("CodeAlive parameter should be logical (column name), or NULL if no presence column.")

  if (!inherits(Paracou6_2016$CircCorr, "numeric") && !is.null(CircCorr)) #DBH
    stop("CircCorr parameter should be numeric (column name), or NULL if no presence column.")

# 
# if (!inherits(Paracou6_2016$Forest, "character"))
#     stop("Forest parameter should be a character (column name), or NULL if no presence column.") #pas besoin
#    if (!inherits(Paracou6_2016$idVern, "integer") && !is.null(idVern))
#     stop("idVern parameter should be an integer (column name), or NULL if no presence column.") #pas besoin
# 
#   
#    
#    if (!inherits(Paracou6_2016$PlotArea, "numeric") && !is.null(PlotArea))
#     stop("PlotArea parameter should be numeric (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$SubPlot, "integer") && !is.null(SubPlot))
#     stop("SubPlot parameter should be an integer (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$TreeFieldNum, "numeric") && !is.null(TreeFieldNum))
#     stop("TreeFieldNum parameter should be numeric (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$Project, "character") && !is.null(Project))
#     stop("Project parameter should be a character (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$Protocole, "character") && !is.null(Protocole))
#     stop("Protocole parameter should be a character (column name), or NULL if no presence column.") #pas besoin
```

```{r all with error message}
if(!all("Plot" %in% names(Paracou6_2016),
"CensusYear" %in% names(Paracou6_2016),
"idTree" %in% names(Paracou6_2016),
"Family" %in% names(Paracou6_2016),
"Genus" %in% names(Paracou6_2016),
"Species" %in% names(Paracou6_2016),
"CircCorr" %in% names(Paracou6_2016),
"CodeAlive" %in% names(Paracou6_2016),
"CommercialSp" %in% names(Paracou6_2016),
"UTMZone" %in% names(Paracou6_2016),
'Lat' %in% names(Paracou6_2016),
'Lon' %in% names(Paracou6_2016),
'VernName' %in% names(Paracou6_2016),
'Xfield' %in% names(Paracou6_2016),
'Yfield' %in% names(Paracou6_2016),
'Xutm' %in% names(Paracou6_2016),
'Yutm' %in% names(Paracou6_2016),
if (!inherits(Paracou6_2016$idTree, "integer") && !is.null(idTree)) #majuscule ?
    stop("idTree parameter should be an integer (column name), or NULL if no presence column."),

if (!inherits(Paracou6_2016$Plot, "character") && !is.null(Plot))
    stop("Plot parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Xfield, "numeric") && !is.null(Xfield))
    stop("Xfield parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Yfield, "numeric") && !is.null(Yfield))
    stop("Yfield parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Xutm, "numeric") && !is.null(Xutm))
    stop("Xutm parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Yutm, "numeric") && !is.null(Yutm))
    stop("Yutm parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$UTMZone, "integer") && !is.null(UTMZone))
    stop("UTMZone parameter should be an integer (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Lat, "numeric") && !is.null(Lat))
    stop("Lat parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Lon, "numeric") && !is.null(Lon))
    stop("Lon parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Family, "character") && !is.null(Family))
    stop("Family parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Genus, "character") && !is.null(Genus))
    stop("Genus parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Species, "character") && !is.null(Species))
    stop("Species parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$VernName, "character") && !is.null(VernName))
    stop("VernName parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$CommercialSp, "logical") && !is.null(CommercialSp))
    stop("CommercialSp parameter should be logical (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$CensusYear, "integer") && !is.null(CensusYear))
    stop("CensusYear parameter should be an integer (column name), or NULL if no presence column."),

  if (!inherits(Paracou6_2016$CodeAlive, "logical") && !is.null(CodeAlive))
    stop("CodeAlive parameter should be logical (column name), or NULL if no presence column."),

  if (!inherits(Paracou6_2016$CircCorr, "numeric") && !is.null(CircCorr)) #DBH
    stop("CircCorr parameter should be numeric (column name), or NULL if no presence column.")))
    stop("Your data format, as specified in the 'data' argument, is not compatible. See the package vignette for the required format.") 
```

>>>>

```{r}
library(dplyr)
library(tibble)
library(tidyr)
```

```{r}
ParamCrownDiameterAllometry <- read_delim("D:/VSC Kourou/DATA/ParamCrownDiameterAllometry.csv", 
    ";", escape_double = FALSE, col_types = cols(alpha = col_number(), 
        beta = col_number()), locale = locale(decimal_mark = ","), 
    trim_ws = TRUE) %>%
  separate(ScientificName, sep = "_", into = c("Genus", "Species"), remove = F) %>% 
  mutate(Family = ifelse(Taxo == "fam", Genus, NA)) %>% 
  mutate(Genus = ifelse(is.na(Family), Genus, NA))

usethis::use_data(ParamCrownDiameterAllometry)
```

# Data loading
```{r}
data(Paracou6_2016) #data inventory in the global environment
data("ForestZoneVolumeParametersTable") # The volume parameters data in the global environment
load(system.file("extdata", "BrokenParacou6_2016.rda", package = "Maria")) # the test inventory in the global environment
load(system.file("extdata", "ParamCrownDiameterAllometry.rda", package = "Maria"))
```
```{r}
lapply(Paracou6_2016, class) #give the class of all the data variables
```

# Create a broken dataset
```{r Create a broken dataset}
#in inst/extdata
library(schoolmath)
library(dplyr)

BrokenParacou6_2016 <- Paracou6_2016 %>% 
  select(-Species) %>% # missing column
  rename(Vernacular = VernName) %>% # other column name
  rename(PLOT = Plot) %>% # variable name in upper case
  mutate(idTree = as.character(as.numeric(idTree))) %>% #an other variable class
  mutate(Lon = round(Lon, digits =0)) %>% # no digits
  mutate(Genus = toupper(Genus)) %>%  # values in upper case
  mutate(CodeAlive = ifelse(is.even(Paracou6_2016$idTree), "FALSE", CodeAlive)) %>% # Some FALSE in CodeAlive
  mutate(CensusYear = ifelse(is.even(Paracou6_2016$idTree), CensusYear, CensusYear+1)) # different years inventory
BrokenParacou6_2016[1:3, "idTree"] <- "200" # create a 2 time present tree
BrokenParacou6_2016[4, "CircCorr"] <- "6" # DBH < 10
BrokenParacou6_2016[5, "Plot"] <- "100" # a tree in another plot

# pour l'enregistrer : usethis::use_data(BrokenParacou6_2016) puis le move dans ins/exdata
```

# inventorycheckformat
INPUT: inventory = Inventaires botaniques Guyafor/ONF (data frame)
OUTPUT: inventory = Inventaires botaniques type “Guyafor”(data frame)

vérification du type d’inventaire : 
* Test présence des variables.
* Test class des variables.
Si les tests précédents ont tous reçu =TRUE, STOP de la fonction inventorycheckformat 
Sinon mettre en forme “inventaire Guyafor” (nom des var, leurs classes, unités (mettre des limites avec range),...)
re-vérification du type d’inventaire : 
* Test présence des variables
* Test class des variables
Si les tests précédent ont tous reçu =TRUE, STOP de la fonction inventorycheckformat
Sinon Message : “Votre format de données renseignés à l’argument “inventory”, n’est pas compatible. Voir la vignette du package pour y trouver le format requis.”

@@

'pouet' %in% names(Paracou6_2016)
if (!is.null(Paracou6_2016$pouet)) 
    stop("pouet variable should be an integer.")

@@

```{r}
# Variables presence check
if(!("Plot" %in% names(Paracou6_2016)))
  stop("Plot variable is not found")
if(!("CensusYear" %in% names(Paracou6_2016)))
    stop("CensusYear variable is not found")
if(!("idTree" %in% names(Paracou6_2016)))
    stop("idTree variable is not found")
if(!("Family" %in% names(Paracou6_2016)))
    stop("Family variable is not found")
if(!("Genus" %in% names(Paracou6_2016)))
    stop("Genus variable is not found")
if(!("Species" %in% names(Paracou6_2016)))
    stop("Species variable is not found")
if(!("CircCorr" %in% names(Paracou6_2016)))
  stop("CircCorr variable is not found")
if(!("CodeAlive" %in% names(Paracou6_2016)))
    stop("CodeAlive variable is not found")
if(!("CommercialSp" %in% names(Paracou6_2016)))
    stop("CommercialSp variable is not found")
if(!("UTMZone" %in% names(Paracou6_2016)))
    stop("UTMZone variable is not found")
if(!('Lat' %in% names(Paracou6_2016)))
    stop("Lat variable is not found")
if(!('Lon' %in% names(Paracou6_2016)))
    stop("Lon variable is not found")
if(!('VernName' %in% names(Paracou6_2016)))
    stop("VernName variable is not found")
if(!('Xfield' %in% names(Paracou6_2016)))
    stop("Xfield variable is not found")
if(!('Yfield' %in% names(Paracou6_2016)))
    stop("Yfield variable is not found")
if(!('Xutm' %in% names(Paracou6_2016)))
    stop("Xutm variable is not found")
if(!('Yutm' %in% names(Paracou6_2016)))
    stop("Yutm variable is not found")

@@

```{r}
# Variables class check
 if (!inherits(Paracou6_2016$idTree, "integer") && !is.null(idTree)) #majuscule ?
    stop("idTree variable should be an integer.")

if (!inherits(Paracou6_2016$Plot, "character") && !is.null(Plot))
    stop("Plot variable should be a character.")

   if (!inherits(Paracou6_2016$Xfield, "numeric") && !is.null(Xfield))
    stop("Xfield variable should be numeric.")

   if (!inherits(Paracou6_2016$Yfield, "numeric") && !is.null(Yfield))
    stop("Yfield variable should be numeric.")

   if (!inherits(Paracou6_2016$Xutm, "numeric") && !is.null(Xutm))
    stop("Xutm variable should be numeric.")

   if (!inherits(Paracou6_2016$Yutm, "numeric") && !is.null(Yutm))
    stop("Yutm variable should be numeric.")

   if (!inherits(Paracou6_2016$UTMZone, "integer") && !is.null(UTMZone))
    stop("UTMZone variable should be an integer.")

   if (!inherits(Paracou6_2016$Lat, "numeric") && !is.null(Lat))
    stop("Lat variable should be numeric.")

   if (!inherits(Paracou6_2016$Lon, "numeric") && !is.null(Lon))
    stop("Lon variable should be numeric.")

   if (!inherits(Paracou6_2016$Family, "character") && !is.null(Family))
    stop("Family variable should be a character.")

   if (!inherits(Paracou6_2016$Genus, "character") && !is.null(Genus))
    stop("Genus variable should be a character.")

   if (!inherits(Paracou6_2016$Species, "character") && !is.null(Species))
    stop("Species variable should be a character.")

   if (!inherits(Paracou6_2016$VernName, "character") && !is.null(VernName))
    stop("VernName variable should be a character.")

   if (!inherits(Paracou6_2016$CommercialSp, "logical") && !is.null(CommercialSp))
    stop("CommercialSp variable should be logical.")

   if (!inherits(Paracou6_2016$CensusYear, "integer") && !is.null(CensusYear))
    stop("CensusYear variable should be an integer.")

  if (!inherits(Paracou6_2016$CodeAlive, "logical") && !is.null(CodeAlive))
    stop("CodeAlive variable should be logical.")

  if (!inherits(Paracou6_2016$CircCorr, "numeric") && !is.null(CircCorr)) #DBH
    stop("CircCorr variable should be numeric.")

# 
# if (!inherits(Paracou6_2016$Forest, "character"))
#     stop("Forest variable should be a character.") #pas besoin
#    if (!inherits(Paracou6_2016$idVern, "integer") && !is.null(idVern))
#     stop("idVern variable should be an integer.") #pas besoin
# 
#   
#    
#    if (!inherits(Paracou6_2016$PlotArea, "numeric") && !is.null(PlotArea))
#     stop("PlotArea variable should be numeric.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$SubPlot, "integer") && !is.null(SubPlot))
#     stop("SubPlot variable should be an integer.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$TreeFieldNum, "numeric") && !is.null(TreeFieldNum))
#     stop("TreeFieldNum variable should be numeric.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$Project, "character") && !is.null(Project))
#     stop("Project variable should be a character.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$Protocole, "character") && !is.null(Protocole))
#     stop("Protocole variable should be a character.") #pas besoin
```

```{r all with error message}
if(!("Plot" %in% names(Paracou6_2016)))
  stop("Plot variable is not found")
if(!("CensusYear" %in% names(Paracou6_2016)))
    stop("CensusYear variable is not found")
if(!("idTree" %in% names(Paracou6_2016)))
    stop("idTree variable is not found")
if(!("Family" %in% names(Paracou6_2016)))
    stop("Family variable is not found")
if(!("Genus" %in% names(Paracou6_2016)))
    stop("Genus variable is not found")
if(!("Species" %in% names(Paracou6_2016)))
    stop("Species variable is not found")
if(!("CircCorr" %in% names(Paracou6_2016)))
  stop("CircCorr variable is not found")
if(!("CodeAlive" %in% names(Paracou6_2016)))
    stop("CodeAlive variable is not found")
if(!("CommercialSp" %in% names(Paracou6_2016)))
    stop("CommercialSp variable is not found")
if(!("UTMZone" %in% names(Paracou6_2016)))
    stop("UTMZone variable is not found")
if(!('Lat' %in% names(Paracou6_2016)))
    stop("Lat variable is not found")
if(!('Lon' %in% names(Paracou6_2016)))
    stop("Lon variable is not found")
if(!('VernName' %in% names(Paracou6_2016)))
    stop("VernName variable is not found")
if(!('Xfield' %in% names(Paracou6_2016)))
    stop("Xfield variable is not found")
if(!('Yfield' %in% names(Paracou6_2016)))
    stop("Yfield variable is not found")
if(!('Xutm' %in% names(Paracou6_2016)))
    stop("Xutm variable is not found")
if(!('Yutm' %in% names(Paracou6_2016)))
    stop("Yutm variable is not found")
if (!inherits(Paracou6_2016$idTree, "integer.")) #majuscule ?
    stop("idTree variable should be an integer.")

if (!inherits(Paracou6_2016$Plot, "character"))
    stop("Plot variable should be a character.")

   if (!inherits(Paracou6_2016$Xfield, "numeric"))
    stop("Xfield variable should be numeric.")

   if (!inherits(Paracou6_2016$Yfield, "numeric"))
    stop("Yfield variable should be numeric.")

   if (!inherits(Paracou6_2016$Xutm, "numeric"))
    stop("Xutm variable should be numeric.")

   if (!inherits(Paracou6_2016$Yutm, "numeric"))
    stop("Yutm variable should be numeric.")

   if (!inherits(Paracou6_2016$UTMZone, "integer"))
    stop("UTMZone variable should be an integer.")

   if (!inherits(Paracou6_2016$Lat, "numeric"))
    stop("Lat variable should be numeric.")

   if (!inherits(Paracou6_2016$Lon, "numeric"))
    stop("Lon variable should be numeric.")

   if (!inherits(Paracou6_2016$Family, "character"))
    stop("Family variable should be a character.")

   if (!inherits(Paracou6_2016$Genus, "character"))
    stop("Genus variable should be a character.")

   if (!inherits(Paracou6_2016$Species, "character"))
    stop("Species variable should be a character.")

   if (!inherits(Paracou6_2016$VernName, "character"))
    stop("VernName variable should be a character.")

   if (!inherits(Paracou6_2016$CommercialSp, "logical"))
    stop("CommercialSp variable should be logical.")

   if (!inherits(Paracou6_2016$CensusYear, "integer"))
    stop("CensusYear variable should be an integer.")

  if (!inherits(Paracou6_2016$CodeAlive, "logical"))
    stop("CodeAlive variable should be logical.")

  if (!inherits(Paracou6_2016$CircCorr, "numeric")) #DBH
    stop("CircCorr variable should be numeric.")
```

# cleaninventory
INPUT: inventory = Inventaires botaniques type “Guyafor” (Paracou6_2016)(data frame)
OUTPUT: inventory = Inventaires botaniques exploitable (Paracou6_2016Clean) (data frame)

Ne prendre que les arbres vivants : lignes ayant la valeur “TRUE” à la colonne “CodeAlive” de inventory. 
Check si les valeurs de ID sont uniques. Si FALSE message d’erreur “Les identifiants des arbres ne sont pas uniques”
Enlever les arbres hors de la parcelle s’il y en a : coordonnées de l’arbre non comprises dans les coordonnées de la parcelle.
Ne prendre que les lignes pour lesquelles “DBH”> 10.
```{r}
duplicated(Paracou6_2016)
unique(Paracou6_2016) #same rows number

#duplicated() gives "FALSE" when the row is unique
# any() : at least one of the values true?
any(duplicated(Paracou6_2016)) # at least one row is duplicated?
all(Paracou6_2016$Plot==Paracou6_2016$Plot[1]) #all the Plot values are equal?
```

```{r}

if (!("DBH" %in% names(Paracou6_2016))) {add_column(Paracou6_2016, DBH = NA) #if DBH doesn't exist create it
   Paracou6_2016$DBH = Paracou6_2016$CircCorr/pi} # and compute it

InventoryClean <- Paracou6_2016 %>% 
  filter(CodeAlive == "TRUE") %>% #only alive trees
  filter(DBH >= 10) # DBH >= 10

  if (any(duplicated(Paracou6_2016)==TRUE))
    stop ("Tree identifiers (idTree) are not unique.") # stop function if the tree identifiers (idTree) are not unique
  if (all(Paracou6_2016$Plot==Paracou6_2016$Plot[1])==FALSE) #all the Plot values are equal?
    stop ("Your inventory concerns different plots (Plot). Our function simulates logging at the plot level.") 
  # Remove out-plot trees (A faire)

# addtreedim
Fonction d’ajout de variables pour l’INPUT “inventory” :
INPUT: inventory = InventoryClean (data frame)
OUTPUT: inventory = Inventory_addtreedim (data frame)

calculer pour tous les arbres : 
  + TreeBiomass : biomasse de tous les arbres (BIOMASS package)
  + TreeHeight : hauteur de l’arbre (H) (BIOMASS package)
  + TreeHarvestableVolume : volume exploitable à partir du tarif de cubage (= a + b*DBH2) que représente chaque arbre (a et b dépendent de la localisation)
  + TrunkHeight : hauteur de fût (TreeHarvestableVolume  = π(DBH/2)² x TrunkHeight)
  + CrownDiameter : diamètre de couronne (CD) (ln(D) = 𝜶+ 𝜷 ln(H*CD) + 𝜺 (allométries de Mélaine)
  + CrownHeight : TreeHeight - TrunkHeight

```{r TreeHeight}
library(BIOMASS)
library(knitr)
# TreeHeight = hauteur de l’arbre (H) (en mètre) (BIOMASS package)
# 3 options :
# - get my tree heights from the location (from continent or region specific model, or from coordinates)
# - from other Guyanese trees heights data, fit different H-D models, select the best and apply on my data.

RegionTreeHeight <- retrieveH(D = Paracou6_2016$CircCor, region = "GuianaShield") #continent or region specific model (Feldpaush et al. (2012))
# GuianaShield = Guyana, French Guiana (piste St Elie (Chave 2005)), Venezuela

n <- length(Paracou6_2016$CircCor)
coord <- cbind(long = rep(-52.92, n), lat = rep(5.27, n)) # Paracou P6 coordinates
CoordTreeHeight <- retrieveH(D = Paracou6_2016$CircCor, coord = coord) #by coordinates (Chave et al. (2012))

result <- modelHD(D = , H = , useWeight = TRUE) #from Guyanese trees heights data
kable(result)
# Compute the local H-D model with the lowest RSE (Root Squared Error)
selectedHDmodel <- modelHD(D = , H = , method = "log2", useWeight = TRUE)

predictHeight(CircCor, model)

```{r TreeHarvestableVolume}
# TreeHarvestableVolume = volume exploitable à partir du tarif de cubage (= a + b*DBH2) que représente chaque arbre (a et b dépendent de la localisation)

ForestZoneVolumeParametersTable <- data.frame (Forest = c("Acarouany","BAFOG","Kaw","Laussat","Montagne Plomb","Montagne Tortue","Nouragues","Organabo","Paracou","Régina St Georges","Risquetout","Tibourou"), # Guyafor forests
                  Zone = c("East", "West", "Central", "FrenchGuiana")) %>% # Volume formules zones
  mutate(Zone = ifelse(Forest == "Acarouany"| Forest =="BAFOG" |Forest =="Laussat", "West", NA)) %>% # West zone
  mutate(Zone = ifelse(Forest == "Montagne Plomb"| Forest =="Organabo"|Forest =="Paracou"|Forest =="Risquetout", "Central", Zone)) %>% # Central zone
  mutate(Zone = ifelse(Forest == "Kaw"| Forest =="Montagne Tortue"|Forest =="Nouragues"|Forest =="Régina St Georges"|Forest =="Tibourou", "East", Zone)) %>% # East zone
  left_join(VolumeParametersTable)

Paracou6_2016_volume <- Paracou6_2016 %>% 
  left_join(ForestZoneVolumeParametersTable) %>% 
mutate(TreeHarvestableVolume = aCoef + bCoef * CircCorr^2) # compute the tree harvestable volume

# West  = c("Acarouany", "BAFOG", "Laussat")
# Central  = c("Montagne Plomb", "Organabo","Paracou", "Risquetout")
# East = c("Kaw", "Montagne Tortue","Nouragues", "Régina St Georges", "Tibourou")
# Guyane = le reste par défaut
# et possibilité à l'utilisateur de mettre ses propres tarifs

```{r TrunkHeight & CrownHeight}
mutate(TrunkHeight = TreeHarvestableVolume/(π(DBH/2)^2)) # compute the trunk height (cylinderVolume= π(DBH/2)² x H)
mutate(CrownHeight = TreeHeight - TrunkHeight) # compute the crown height

```{r}
spParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at species scale
  filter(Taxo == "sp") %>% 
  select(-Family)
genParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at genus scale
  filter(Taxo == "gen") %>% 
  select(-Family, -ScientificName, -Species)
famParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at family scale
  filter(Taxo == "fam") %>% 
  select(-Genus, -ScientificName, -Species)

Paracou6_2016 %>% 
  sample_n(100) %>% # short data
  select(idTree, Family, Genus, Species) %>% #only useful columns 
  left_join(select(spParamCrownDiameter, -Taxo, -ScientificName, -beta), by = c("Genus","Species")) %>% 
  left_join(select(genParamCrownDiameter, -Taxo, -beta), by = c("Genus"), suffix = c(".species", ".genus")) %>% 
  left_join(select(famParamCrownDiameter, -Taxo, -beta) %>% 
              rename(alpha.family = alpha), by = c("Family")) %>% 
  mutate(alpha = alpha.species) %>% #mutate(alpha = alpha.species,beta = beta.species)
  mutate(alpha = ifelse(is.na(alpha), alpha.genus, alpha)) %>% 
  mutate(alpha = ifelse(is.na(alpha), alpha.family, alpha)) %>% 
  mutate(alpha = ifelse(is.na(alpha), mean(alpha, na.rm =T), alpha)) %>% 
  select(-alpha.species, -alpha.genus, -alpha.family)
```

```{r CrownDiameter}

spParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at species scale
  filter(Taxo == "sp") %>% 
  select(-Family)
genParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at genus scale
  filter(Taxo == "gen") %>% 
  select(-Family, -ScientificName, -Species)
famParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at family scale
  filter(Taxo == "fam") %>% 
  select(-Genus, -ScientificName, -Species)

error <- rnorm(2000, mean = 0, sd = 0.0295966977) #Mélaine's model error distribution (hist(error))

Paracou6_2016_crown <- Paracou6_2016 %>% 
  unite(Genus, Species, col = "ScientificName", sep = "_", remove = F) %>% 
  left_join(spParamCrownDiameter, by = c("ScientificName","Genus","Species")) %>% #Add parameters at species scale
  left_join(genParamCrownDiameter, by = "Genus", suffix = c(".species", ".genus")) %>% #Add parameters at genus scale
  left_join(famParamCrownDiameter, by = "Family") %>% #Add parameters at family scale

  rename(Taxo.family = Taxo, alpha.family = alpha, beta.family = beta) %>% 
  mutate(alpha = alpha.species, beta = beta.species, Taxo = Taxo.species) %>% # create the future & unique alpha,beta,taxo columns 
  mutate(alpha = ifelse(is.na(alpha), alpha.genus, alpha)) %>% #if species parameters are absent, take genus parameters
  mutate(alpha = ifelse(is.na(alpha), alpha.family, alpha)) %>% #if species&genus parameters are absent, take family parameters
  mutate(alpha = ifelse(is.na(alpha), mean(alpha, na.rm =T), alpha)) %>% #if species,genus&family parameters are absent, take parameters mean
  select(-alpha.species, -alpha.genus, -alpha.family) %>% #remove obsolete columns

  mutate(beta = ifelse(is.na(beta), beta.genus, beta)) %>% #if species parameters are absent, take genus parameters
  mutate(beta = ifelse(is.na(beta), beta.family, beta)) %>% #if species&genus parameters are absent, take family parameters
  mutate(beta = ifelse(is.na(beta), mean(beta, na.rm =T), beta)) %>% #if species,genus&family parameters are absent, take parameters mean
  select(-beta.species, -beta.genus, -beta.family) %>% #remove obsolete columns

  mutate(Taxo = ifelse(is.na(Taxo), Taxo.genus, Taxo)) %>% #if species parameters are absent, take genus parameters
  mutate(Taxo = ifelse(is.na(Taxo), Taxo.family, Taxo)) %>% #if species&genus parameters are absent, take family parameters
  mutate(Taxo = ifelse(is.na(Taxo), "mean", Taxo)) %>% #if species,genus&family parameters are absent, take parameters mean
  select(-Taxo.species, -Taxo.genus, -Taxo.family) %>%  #remove obsolete columns
  mutate(CrownDiameter = exp((
  (ln(DBH)-alpha-error)/beta)
  )/TreeHeight) # compute the crown diameter (CD) (ln(D) = 𝜶+ 𝜷 ln(H*CD) + 𝜺, with 𝜺~N(0,σ^2) and meanσ^2 = 0.0295966977. (Mélaine's allometries))

# Informations:
# How many individuals have find their Crown diameter parameters at species, genus, family or no scale:
sum(Paracou6_2016_crown$Taxo == "fam") #442 individuals with family scale parameters
sum(Paracou6_2016_crown$Taxo == "gen") #742 individuals with genus scale parameters
sum(Paracou6_2016_crown$Taxo == "sp") #2364 individuals with family scale parameters
sum(Paracou6_2016_crown$Taxo == "mean") #72 individuals no recognize at any scale, so with plot mean parameters

unique(Paracou6_2016_crown[(Paracou6_2016_crown$Taxo == "mean"),]$ScientificName) #which species are no recognized
length(unique(Paracou6_2016_crown$ScientificName)) #How many species are in the plot
```

```{r All addtreedim}
# allometry parameters data preparation:
spParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at species scale
  filter(Taxo == "sp") %>% 
  select(-Family)
genParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at genus scale
  filter(Taxo == "gen") %>% 
  select(-Family, -ScientificName, -Species)
famParamCrownDiameter <- ParamCrownDiameterAllometry %>% #parameters at family scale
  filter(Taxo == "fam") %>% 
  select(-Genus, -ScientificName, -Species)

error <- rnorm(2000, mean = 0, sd = 0.0295966977) #Mélaine's model error distribution (hist(error))

# Variables computation:
InventoryAddtreedim <- InventoryClean %>%

# TreeHarvestableVolume
  left_join(ForestZoneVolumeParametersTable) %>% 
  mutate(TreeHarvestableVolume = aCoef + bCoef * CircCorr^2) %>%  # compute the tree harvestable volume

# TrunkHeight & CrownHeight  
  mutate(TrunkHeight = TreeHarvestableVolume/(pi*(DBH/2)^2)) %>%  # compute the trunk height (cylinderVolume= π(DBH/2)² x H)
  # mutate(CrownHeight = TreeHeight - TrunkHeight) %>%  # compute the crown height TREEHEIGHT EN ATTENTE

# CrownDiameter
  unite(Genus, Species, col = "ScientificName", sep = "_", remove = F) %>% 
  left_join(spParamCrownDiameter, by = c("ScientificName","Genus","Species")) %>% #Add parameters at species scale
  left_join(genParamCrownDiameter, by = "Genus", suffix = c(".species", ".genus")) %>% #Add parameters at genus scale
  left_join(famParamCrownDiameter, by = "Family") %>% #Add parameters at family scale
